---
title: "p8105_hw2_sq2266"
author: "Sihan Qiu"
output: github_document
---




```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1
```{r}
transit_df = 
  read_csv("NYC_Transit_Subway_Entrance_And_Exit_Data.csv",na = c("NA",".","")) |>
  janitor::clean_names() |>
  mutate(entry = case_match(
    entry,
    "YES" ~ "TRUE",
    "NO" ~ "FALSE"
  )
  )
print(transit_df)
```

## I am trying to combine key variables
```{r}
    select(transit_df,line,station_name, station_latitude, route1,route2,route3,route4,route5,route6,route7,route8,route9,route10, entry, vending, entrance_type, ada) |>

print(transit_df)
```


Description: 
This dataset includes information of subway stations in New York City. Key variables includes `division`,`line`, `station_name`, `station_latitude`,`station_longitude`,`routes`, `entry`,`vending`,`entrance_type`, and `ada`. The data cleaning steps includes standardizing column names by using `janitor::clean_names` to make uppercase and lowercase separated, converting `entry` from a character variable to a logical variable. 




## Problem 2
### cleaning and reading data
```{r}
mr_trash_df = 
  read_excel("202409 Trash Wheel Collection Data.new.xlsx", sheet = 1, na = c("NA",".","")) |>
  janitor::clean_names() |>
  select(-x15, -x16) |>
  mutate(sports_balls = as.integer(round(sports_balls,0))) |>
  mutate(trash_wheel = "Mr. Trash Wheel") |>
  relocate(trash_wheel)
```


### cleaning Professor Trash Wheel and Gwynnda
```{r}
prof_trash_df = 
  read_excel("202309 Trash Wheel Collection Data.xlsx", sheet = 2, na = c("NA",".","")) |>
  janitor::clean_names() |>
  mutate(trash_wheel = "Profesor Trash Wheel") |>
  relocate(trash_wheel)

Gwy_trash_df = 
  read_excel("202309 Trash Wheel Collection Data.xlsx", sheet = 4,na = c("NA",".","")) |>
  janitor::clean_names() |>
  mutate(trash_wheel = "Gwynnda") |>
  relocate(trash_wheel)
```

### Combining
```{r}
total_trash_df = 
     left_join(mr_trash_df,prof_trash_df,Gwy_trash_df, by = "trash_wheel")
```


```{r}
Description:
  

```



## Problem 3
create a single, well-organized dataset with all the information contained in these data files
```{r}
bakers_df = 
  read_csv("gbb_datasets/bakers.csv", na = c("NA",".",""))|>
  janitor::clean_names()|>
  rename(baker = baker_name)
print(bakers_df)
```

```{r}
bakes_df = 
  read_csv("gbb_datasets/bakes.csv",na = c("NA",".",""))
  janitor::clean_names()
print(bakes_df)
```


```{r}
results_df = 
  read_csv("gbb_datasets/results.csv",na = c("NA",".","")) |>
  janitor::clean_names() |>
  rename(
    result = in_stayed_in_out_eliminated_star_baker_star_baker_winner_series_winner_runner_up_series_runner_up_wd_withdrew
    
  )
  
print(results_df)
```

## combining datasets
```{r}
bakers_bakes_df = 
  left_join(bakers_df, bakes_df, by = "baker")
print(bakers_bakes_df)
```

```{r}
merged_df = 
  left_join(bakers_bakes_df, results_df, by = "baker") |>
  arrange(series) |>
  relocate(series)
print(merged_df)
```


